Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2021-01-30 paper/paper.tex
paper: Add Conclusions

Normally, the characterization process takes into account astrometric features
such as proper motion in right ascension and declination, and parallax.
Photometric features like $E(b-r) / G_{mag}$ are considered too.
These last variables are used for generating the H-R diagram of the OC candidate stars.
This stars should present a sharp profile corresponding to an isochrone curve derived
from a theoretical model.
This model does consider metallicity, mass and brightness of the stars involved.
At the moment, a set of tools have been used for this purpose, as explained later,
but in general, these tools require supervised and parameterized models.
So a previous knowledge of the cluster is necessary or the process must be repeated
iteratively in order to achieve valid results.

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
    \end{subfigure}
  \end{subfigure}
  \medskip
  \begin{subfigure}{\columnwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
      \caption{N = 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
      \caption{N = 5}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
      \caption{N = 8}
    \end{subfigure}
  \end{subfigure}
  \caption{K-Means comparisons with Melotte 22}
  \label{fig:kmeans_comparisons_melotte_22}
\end{figure}

@@

This is due to the fact that OC's stars are surrounded by other stars with possibly similar
properties. So, setting the number of clusters to two is too low to separate them properly.

@@

Resumen de las contribuciones del trabajo, en el que relaciones las contribuciones y los resultados obtenidos con los objetivos que habías planteado para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.
Finalmente, hablar de líneas de trabajo futuro que podrían aportar valor añadido al TFM realizado. La sección debería señalar las perspectivas de futuro que abre el trabajo desarrollado para el campo de estudio definido. En el fondo, debes justificar de qué modo puede emplearse la aportación que has desarrollado y en qué campos.

\appendix
\section{APPENDIX}
Apéndices, en caso de ser necesario.

>>>>

Normally, the characterization process takes into account astrometric features
such as proper motion in right ascension and declination, and parallax.
Photometric features like $E(b-r) / G_{mag}$ are considered too.
These last variables are used for generating the H-R diagram of the OC candidate stars.
This stars should present a sharp profile corresponding to an isochrone curve derived
from a theoretical model.
This model does consider metallicity, mass and brightness of the stars involved.
At the moment, a set of tools have been used for this purpose, as explained later.
But in general, these tools require supervised and parameterized models.
Therefore, a previous knowledge of the cluster is necessary or the process must be
repeated iteratively in order to achieve valid results.

@@

@@

This is due to the fact that OC's stars are surrounded by other stars with possibly similar
properties. Therefore, setting the number of clusters to two is too low to separate them properly.

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_pm_melotte_22.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_pm_melotte_22.png}
    \end{subfigure}
  \end{subfigure}
  \medskip
  \begin{subfigure}{\columnwidth}
    \centering
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n2_parallax_melotte_22.png}
      \caption{N = 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n5_parallax_melotte_22.png}
      \caption{N = 5}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/kmeans/kmeans_n8_parallax_melotte_22.png}
      \caption{N = 8}
    \end{subfigure}
  \end{subfigure}
  \caption{K-Means comparisons with Melotte 22}
  \label{fig:kmeans_comparisons_melotte_22}
\end{figure}

@@

From the results shown in Section \ref{sec:results}, we can say that, in general,
our model works fine identifying and characterizing open clusters.
So we have succeeded building a model non-parameterized and non-supervised
for open cluster characterization. However, some improvements should be done
in order to improve its accuracy and precision.

We have built a non-parameterized model since the user is not required to provide
any information about the cluster profile or its properties.
Nevertheless some tweaks can be done regarding the cluster, like changing the initial
number of clusters or modifying the clustering layer structure in order to improve
the final characterization.

Another key point in our model is the initializer kernel.
This kernel prepares data before passing it to the ANN
and the results may vary significantly depending on this kernel.
This is something to avoid, so an improvement to solve this issue is necessary
in order to have a reliable model which does not depend on the dataset order.

Despite that, comparing our method with Clusterix, we can still say that the
number of parameters (or hyperparameters) in our model is smaller and that
we do not need previous knowledge about the studied cluster.
That makes easier to test different configurations and automate the process.

\subsection{FURTHER RESEARCH}

One open point to do from now is testing the model with a wider range of clusters.
In general we could think of applying it to the whole VizieR catalogue.
That way we could have a better idea about the current limitations of our model
and we could make some adjustments to improve it. We could determine with higher
accuracy the typologies that our model works well with, and which ones the model
does not resolve properly. Furthermore, we could stablish different sets of
hyperparameters regarding the typology of the cluster.

With our model, new uncatalogued clusters arise from the data apart of the main ones.
Hence, it is also interesting the individual study of these new clusters.

Another possible change in our model would be using DBSCAN instead of KMeans
as our initial clustering algorithm. Maybe this algorithm gives us a better starting
point for the DEC model which could improve the results.

Finally, we have used the Gaia DR2 database as our data source for this work,
but recently the DR3 dataset has been released.
Therefore, it is evident the interest on testing our model with this new data.
